# Practical Machine Learning - Prediction Assignment
# --------------------------------------------------

### For this assignment I analyzed the provided data to determine what activity an individual performs. Packages "caret" and "randomForest" were utilized to get correct results. Seed value was used for consistent result


```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(2048)
options(warn=-1)
```

### Data for test and training was loaded into respective data frames
### Some values which contained "#DIV/0!" were replaced with an NA value

```{r}
training_data <- read.csv("pml-training.csv",sep=",",na.strings = c("NA",""),header=TRUE)
evaluation_data <- read.csv("pml-testing.csv", sep=",",na.strings = c("NA",""),header=TRUE)
```

### All columns 8 were assigned to be numeric

```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

### Some columns were mostly blank.  These did not contribute well to the prediction.  I chose a feature included complete columns.  Removed user name, timestamps and windows.  
### Determine and display out feature set

```{r}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

### Model data was built from feature set

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

### Build 5 random forests with 150 trees each. Use parallel processing to build this model
### model. I found several examples of how to perform parallel processing with random forests in R, this


```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

### Provide error reports for both training and test data
```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

# Conclusions and Test Data Submit
# --------------------------------

### Based on the confusion matrix this model is accurate
### Submission of the final code for problem file generation

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```